---
title: "Prediction Assignment: Personal Weight Lifting Exercise"
author: "Trunojoyo Anggara"
date: "February 20, 2016"
output: html_document
---

# Executive Summary

Linear Discriminant Analysis (LDA) and Random Forest were used to predict which activities were being performed in the testing dataset from [Human Activity Recognition](http://groupware.les.inf.puc-rio.br/har). 

The model were built with 10-fold cross-validation resampling. The accuracy for LDA and Random Forest are ~0.85 and 0.99, respectively. Seeing the 99% accuracy in random forest model, I used this fit model to predict the 20 different test cases given by the project. 

# Analysis

## Reading Data 

```{r cache=TRUE}
trainingset<-'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
testset<-'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

if (!file.exists('pml-training.csv')) {download.file(trainingset,destfile='pml-training.csv') }
if (!file.exists('pml-testing.csv')) {download.file(testset,destfile='pml-testing.csv')}
training<-read.csv('pml-training.csv')
testing<-read.csv('pml-testing.csv')
```

## Data Partition and Data Cleaning

```{r, cache=TRUE}
# Create Data Partition with 0.7 as Train1 and 0.3 as Test1
library(caret)
inTrain<-createDataPartition(y=training$classe, p=0.7, list=FALSE)
Train1<-training[inTrain,]
Test1<-training[-inTrain,]
```

Let's clean the "Train1" dataset:
```{r, cache=TRUE}
# There are 160 variables, but some of them have identical values 
# and cannot be used for predictor
# We will use nearZeroVar to identify them
# If I want to know the details: 
# nearZeroVar(Train1,saveMetrics = TRUE)
DelCol<-nearZeroVar(Train1)
Trainmod<-Train1[,-DelCol]
dim(Trainmod)
```

This reduces the variables into 108 variables. Let's do more cleaning:

```{r}
# Delete the X: row number variable
Trainmod<-Trainmod[-1]
# names(Trainmod)

# Some of the columns still have a lot of NAs that we can take away from the 
# dataset. Using the technique taken from:
# http://stackoverflow.com/questions/10574061/show-columns-with-nas-in-a-data-frame

nacols <- function(df) {
  colnames(df)[unlist(lapply(df, function(x) any(is.na(x))))]
}

ColwithNa<-nacols(Trainmod)
# head(ColwithNa)
# Truncate the Trainmod dataset to not include columns that have NAs 

Trainmod<-Trainmod[,!(names(Trainmod) %in% ColwithNa)]
dim(Trainmod)

# Apply them to the testing data frame 

# Since the last dataset that we want to predict, "testing"
# does not have "classe", so we take the classe column away
ColNameFinal<-colnames(Trainmod)
NoClasse<-ColNameFinal[-58]

Testmod<-Test1[ColNameFinal]
testingmod<-testing[NoClasse]
```

## Linear Discriminant Analysis and Random Forest Algorithm

Let us run the Linear Discriminant Analysis and Random Forest algorithm with 10-fold cross validation resampling method.
```{r, cache=TRUE}
# Linear Discriminant Analysis
library(parallel)
library(doParallel)
set.seed(12345)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv",
                           number = 10,
                           allowParallel = TRUE)

fitlda <- train(classe~., method="lda",data=Trainmod,trControl = fitControl)


# fitlda$finalModel
predictlda<-predict(fitlda,Testmod)
confusionMatrix(predictlda,Testmod$classe)

# Random Forest algorithm
set.seed(12345)
registerDoParallel(cluster)

fitrf <- train(classe~., method="rf",data=Trainmod,trControl = fitControl)
stopCluster(cluster)

#fitrf$finalModel
predictrf<-predict(fitrf,Testmod)
confusionMatrix(predictrf,Testmod$classe)
```
As we see above, the accuracy for LDA is ~0.85. However, the random forest algorithm is ~0.99. Since random forest method have higher accuracy, we will use the random forest fitting to predict the "testingmod"" dataset.

## Out of Sample Error

The out of sample error is 1-accuracy. For LDA, the error is 1-0.85 = 0.15, and for Random Forest is 1-0.99 = 0.01.

## Prediction

Here are the prediction:
```{r, cache=TRUE}
predictreal<-predict(fitrf,testingmod)
predictreal
```
